---
title: "Social Media"
author: "Shubham Bhargava"
date: "2024-04-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(FactoMineR)
library(factoextra)

social_media <- read_excel("social_media_cleaned.xlsx")


### Discriminant Analysis

```{r}
library(MASS)
library(readxl)
library(ROCR)

mydata <- read_excel("social_media_cleaned.xlsx")
mydata$Binary_tfs <- ifelse(mydata$Trouble_falling_asleep == "1", 1, 0)
```

#### Model Development

```{r}
lda_model <- lda(Binary_tfs ~ Instagram +	LinkedIn + SnapChat + Twitter +	`Whatsapp/Wechat` +	youtube +	OTT +	Reddit, data = mydata)
```

#### Model Acceptance

```{r}
summary(lda_model)
print(lda_model)
```

#### Residual Analysis

```{r}
plot(lda_model)
```

#### Prediction
```{r}
lda_predictions <- predict(lda_model, newdata = mydata)
lda_predictions

predicted_classes <- lda_predictions$class
predicted_classes
lda_predictions$x

predicted_probabilities <- as.data.frame(lda_predictions$posterior)
predicted_probabilities
pred <- prediction(predicted_probabilities[,2], mydata$Binary_tfs)
```

#### Model Accuracy

```{r}
roc.perf = performance(pred, measure = "tpr", x.measure = "fpr")
auc.train <- performance(pred, measure = "auc")
auc.train <- auc.train@y.values
plot(roc.perf, main = "ROC Curve", col = "purple", lwd = 2)
abline(a = 0, b = 1, lty = 2, col = "blue")
text(x = .25, y = .65 ,paste("AUC = ", round(auc.train[[1]],3), sep = ""))
```