---
title: "Homework 9"
author: "Shubham Bhargava"
date: "2024-04-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment 9
### Discriminant Analysis

```{r}
library(MASS)
library(readxl)
library(ROCR)

mydata <- read_excel("News Website Dataset_2.xlsx")
mydata$Binary_Content_Category <- ifelse(mydata$Content_Category == "Homepage", 1, 0)

```

#### Model Development
<p>With the use of linear discriminant analysis (LDA), this code trains a model to predict the Binary_Content_Category variable. To make these predictions, it makes use of a number of predictor factors, including Total_Sessions, Conversion_Rate, Avg_Session_Duration, Total Revenue, Content_Category_Num, and Device_Category_Num.</p>

```{r}
lda_model <- lda(Binary_Content_Category ~ Total_Sessions + Conversion_Rate + Avg_Session_Duration + Total_revenue + Content_Category_Num + Device_Category_Num , data = mydata)

```

#### Model Acceptance

<p>The trained Linear Discriminant Analysis (LDA) model, designated as lda_model, is reviewed by the code and approved. It creates a comprehensive summary of the model's parameters using the summary function, encompassing information such as the prior probabilities for every category, the average values for every group, and the coefficients that were employed in the classification procedure. Understanding the characteristics and performance of the model is aided by this synopsis. Furthermore, the print function provides a more thorough examination of the model by displaying other data like group means, coefficients, and classification rules, all of which improve our comprehension of the model's workings. We may totally assess and approve the LDA model by utilizing these capabilities, guaranteeing its application will be clear and dependable.</p>

```{r}
summary(lda_model)
print(lda_model)
```

<p>The result offers comprehensive details on several facets of the Linear Discriminant Analysis (LDA) model. It contains important information such as the counts of observations in each category, the average values of predictor variables for each category, the probabilities given to each category, and the scaling used for linear discriminants. It also includes a summary of the categories under analysis, numerical values that indicate how the predictors break down, and other pertinent details like the total number of observations and the particular formula and function that were used to build the model. All things considered, these elements aid in comprehending the structure of the model and how it uses the data to generate predictions.</p>

<p>The output provides a summary of an LDA (linear discriminant analysis) model's findings. It illustrates how the model uses a variety of input parameters, including Total_Sessions, Conversion_Rate, Avg_Session_Duration, Total_Revenue, Content Category_Num, and Device_Category_Num, to predict a particular outcome (Binary Content Category). The probability of each outcome group (0 and 1) happening is also included in the summary, showing how frequent each category is in the data. It also gives the average input factor values for every output group, which can help identify any discrepancies between the groups. Furthermore, the strength with which each input element affects the model's prediction is indicated by the coefficients of the linear discriminants (LD1). All things considered, this data facilitates comprehension of the model's operation and the important variables that influence its decision-making.</p>

#### Residual Analysis
```{r}
plot(lda_model)
```

#### Prediction
```{r}
lda_predictions <- predict(lda_model, newdata = mydata)
lda_predictions

predicted_classes <- lda_predictions$class
predicted_classes
lda_predictions$x

predicted_probabilities <- as.data.frame(lda_predictions$posterior)
predicted_probabilities
pred <- prediction(predicted_probabilities[,2], mydata$Binary_Content_Category)
```

#### Model Accuracy
```{r}
roc.perf = performance(pred, measure = "tpr", x.measure = "fpr")
auc.train <- performance(pred, measure = "auc")
auc.train <- auc.train@y.values
plot(roc.perf, main = "ROC Curve", col = "green", lwd = 2)
abline(a = 0, b = 1, lty = 2, col = "yellow")
text(x = .25, y = .65 ,paste("AUC = ", round(auc.train[[1]],3), sep = ""))

```